---
title: "ComputerIntensiveMethodsPractice"
author: "Olusoji Oluwafemi Daniel and Owokomo Olajumoke Evangelina"
date: "21 September 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(tidyr)
library(purrr)
library(magrittr)
library(dplyr)
```

## Summary

All codes and write ups here are done to practice the concepts taught in Computer Intensive Methods and also to refine some of the codes used in the course using the `purrr` package.

## Sampling Distributions

```{r}

```

## Constructing The Empirical Distribution

The empirical distribution function is a function that assigns the probability $\frac{1}{n}$ to each sample value. It can be used to approximate the distribution function of the random variable in question.

### Using a Generated Sequnece

```{r}
#generate ssome random sequence
x <- sort(seq(-5,5,length=1000))
#probability of each value in x
dx <- dnorm(x)
#Density plot of x
plot(x,dx,type='l')
#Probability Distribution
px <- pnorm(x)
#plot of the Probability Distribution
plot(x,px,type='l')
```

### Using Simulated Values From a Normal Distribution

```{r}
#simulate 1000 values from the normal distribution and sort them
nv <- sort(rnorm(1000,0,1))
# Obtaining the Distribution Function
nvp <- pnorm(nv)
#plotting the distribution function
plot(nv,nvp,type='l')
##getting the empirical distribution
emnv <- (1:length(nv))/length(nv)
lines(nv,emnv,col=3,lwd=2)
```

It is seen that the empirical distribution properly approximates the distribution function even though it is a discrete distribution that assigns an equal probability $\frac{1}{n}$ to all samples.

## Bootstrapping


```{r}
# A sample of 10 shoe sizes
ss <- c(11.201, 10.035, 11.118, 9.055, 9.434, 9.663, 10.403, 11.662, 9.285,8.84)
#summary statistics of shoe size
mean(ss)
var(ss)
sd(ss)
#variacne of the sample mean
var(ss)/length(ss)
#standard error 
sd(ss)/sqrt(length(ss))
```


























